Model validation plots

### 2021-08-13 review/Response to reviews.Rmd

##  {#R1-2}

> In the intro and discussion, I think it would also be helpful to describe a little bit more about mechanisms for why extreme precipitation matters more in fragments compared to continuous forest.
> I wondered if the authors thought this was too obvious (edge effects, etc), but I think it'd be helpful to spell this out more clearly.
> L466 -- 467 offer a new hypothesis, but briefly revisiting the 'old' evidence/hypotheses also seems important.

##  {#R1-3}

> One challenge of using a new modeling approach is making the methods of the paper understandable and to stand alone.
> As written in the methods and results, I found the implementation very difficult to follow.
> A slightly longer summary of the methods would help.

...see also our response to \@ref(R3-3)

> In the results, I was surprised to see only a few p-values and not much else about the models in terms of how they were assessed for goodness of fit.
> While I am not at all suggesting that p-values are the solution, I would have liked a little more sense of how well the models fit.
> One way to do that might be to show the raw data plotted with the model fits so the reader can see, for example, what kind of functions were drawn.
> Another might be to present something in table format (the comparable version of a table of AICs or of the results of frequentist tests).

<!--# make a table I guess? -->

@@

> Line 323-325: I don't know why the growth of plants in fragments would affect the size of plant in CF.

##  {#R2-10}

> Lines 355-359: The analysis did not consider testing a statistical difference between habitat type (see comment 7), so I think this sentence is misleading (overinterpreting the results).

<!--# need to check the line to figure out the source of confusion, but we did test this with 84%CI overlap -->

@@

##  {#R2-11}

> Line 370: There are no confidence intervals in Fig 2, so we don't know if the difference is significant.

<!--# this is not a plot that can have CIs.  The line doesn't imply statistical significance either.  Not sure how to respond -->

@@

> Lines 383 ff: see comment 7

##  {#R2-13}

> Line 412: I think it would be very informative if you were to expand on this and explain how exactly your method differs from what others have used to detect delayed effects.

...see also our response to \@ref(R1-5).
<!--# R1-5 is asking for more discussion on limitations of DLNMs -->

@@

> Lines 451-453: I think herbivory is this is very different from drought.
> Drought reduced nutrient availability while herbivory removes some plant biomass triggering compensatory growth.

##  {#R2-20}

> Lines 468 ff: I suggest deleting this paragraph.
> It doesn't really matter when exactly a plant dies during a growing season, all that matters for population dynamics is how many plants die from one growing season to the next.
> Distributed Lag Non-Linear Models account for the exact timing of the weather events.

<!--# not sure how to respond.  I think there must be some underlying confusion.  Other reviewers like this explanation -->

@@

We moved the explanation of covariates and random effects earlier as suggested.

##  {#R3-4}

> Line 205 - I'm not sure "bimodal" is correct here -- a cubic function can have 2 local extrema, is that what you mean?

<!--# yes, but might change number of knots, so might require more work revising this -->

@@

> Line 205-207 -- "number of knots not important ..." yes, but package mgcv includes tools for checking those assumptions, would be helpful to include residual analysis and the results of gam.check() in an appendix.

##  {#R3-6}

> Line 211 - random effect -- how did you deal with this in the predictions?
> set to zero in design matrix, Choose a particular value, sample from the posterior?
> The code would tell me but not available.

What we show in the figures are not predictions (i.e. results of `predict()` in R), but partial effects plots generated with the `gratia` package (equivalent to `mgcv::plot.gam()`).
For figures 4, 5, and 6 we have added the model intercepts to evaluated smooths and back-transformed (when applicable) the estimates.
These plots, therefore, do not include the additive effects of other smooths, parametric covariates, or random effects.
We think that this is an intuitive way to display the results.
We added a paragraph to the methods explaining our strategy for visualizing the GAM results.

<!--# see also response to comments where we decide to backtransform univariate smooths -->

### 2021-08-13 review/Response to reviews.Rmd

<!--# make a table I guess? -->

##  {#R1-4}

> Related to above, I think it's important to see more of the raw data than are presented in Fig.
> 1.
> (I'm thinking of a version of Fig 3 that shows size in t+1, survival, and flowering rather than effect size on the y-axis, with the data and lines showing the shape of the GAMs).
> This could be an appendix, but it would help with interpreting Figures 3 -- 6.

One way we approached this is by excluding areas of the heatmap that are far from data, equivalent to not drawing regression lines beyond the range of data.
We implemented this with the `exclude.too.far()` function in the `mgcv` package.
Displaying raw data or residuals on the heatmaps was too messy and not informative in a 2d plot.
Instead, we have explored "slices" of these heatmaps in a supplement <!--# TODO, make that supplement! -->.

### 2021-08-16 doc/validate_models.Rmd
switch to k.check() for no plotting

```{r eval=TRUE}
pdf(NULL)
gam.check(s_1ha)
dev.off()
```

@@

```{r eval=TRUE}
# looking for near zero edf
gam(s_1ha_res ~ s(log_size_prev, k = 10, bs = "cs"), gamma = 1.4,
    data = model.frame(s_1ha)) #fine

gam(s_1ha_res ~ s(spei_history, L,
      bs = "cb",
      k = c(3, 35),
      xt = list(bs = "cr")),
    gamma = 1.4,
    select = TRUE,
    data = model.frame(s_1ha)) #fine
```

@@

```{r}
pdf(NULL)
gam.check(s_cf)
dev.off()
```

@@

```{r}
pdf(NULL)
gam.check(g_1ha)
dev.off()
```

@@

```{r}
pdf(NULL)
gam.check(g_cf)
dev.off()
```

@@

```{r}
pdf(NULL)
gam.check(f_1ha)
dev.off()
```

@@

```{r}
pdf(NULL)
gam.check(f_cf)
dev.off()
```

>>>>

```{r eval=TRUE}
k.check(s_1ha)
```

@@

```{r eval=TRUE}
# looking for near zero edf
gam(s_1ha_res ~ s(log_size_prev, k = 10, bs = "cs"), gamma = 1.4,
    data = model.frame(s_1ha)) %>%
  edf() #fine

gam(s_1ha_res ~ s(spei_history, L,
      bs = "cb",
      k = c(3, 35),
      xt = list(bs = "cr")),
    gamma = 1.4,
    select = TRUE,
    data = model.frame(s_1ha)) %>%
  edf()#fine
```

@@

```{r}
k.check(s_cf)
```

@@

```{r}
k.check(g_1ha)
```

@@

```{r}
k.check(g_cf)
```

@@

```{r}
k.check(f_1ha)
```

@@

```{r}
k.check(f_cf)
```
